---
title: "Cluster analysis with all proteomic data"
format: html
editor: visual
---

# Clustering analysis with all proteomics data

## Hierachial clustering visualised with dendrograms

To investigate if the clinically assigned pam50 categories display in the proteomics dataset, we also wanted to do a hierachial clustering analysis. This was visualised using dendrograms.

First, packages and data are loaded.

```{r}
#| echo: false
#| label: load packages and data

# Packages
library(tidyverse)
library(ggdendro)
source("99_proj_func.R")

# data
clinical_data <- read.table("../data/clinical_data.tsv", sep = "\t", header = T)
cm <- read.table("../data/count_matrix_1_77.tsv", header = T) |>
  pivot_longer(cols = -id, names_to = "id2", values_to = "value")|> # Transforming to pivot_longer
  pivot_wider(names_from = id, values_from = value) |>
  column_to_rownames("id2")

```

### Analysis based on entire proteomics data set

Here, the hierachial clustering is done and plotted as a dendrogram using the `hc_dendro_plot` function from the `99_proj_func.R` file. The function first takes the count matrix and calculates the hierarchical clustering. Hereafter, it plots the results as a dendrogram while colouring for pam50-subtypes

```{r}
#| label: Hierachial clustering 
ddg_plot <- hc_dendro_plot(count_matrix = cm, data_type = "full proteomics dataset")
ddg_plot + 
  labs(colour = "pam50")

# save dendrogram plot
ggsave("../results/hc_dendro_all.png", ddg_plot, height = 5, width = 6)
```

**Conclusion**: From the colored labels, it is apparent, that the samples don't group by pam50 categories when a hierachial clustering is made on the entire proteomics dataset.

```{r}
#quarto::quarto_render("05_analysis_clustering.qmd",output_format = "html",execute = T)
#file.rename(from = "05_analysis_clustering.html",
#            to = "../results/05_analysis_clustering.html")
```

```{r}
rm(list = ls())
```
