---
title: "08_correlation"
format: html
editor: visual
---
```{r}
library(tidyverse)
library(corrr)
library(ggcorrplot)
library(ggplot2)
library(reshape)
library(purrr)
library(rcompanion)
library(patchwork)
```


## Correlation of 4 subgroups with ER, PR and HER2

Load clinical data 
```{r}
clinical_data <- read.table("../data/clinical_data.tsv", sep = "\t", header = T)
```


One-out-of-k encoding for the 4 subgroups
```{r}
clinical_data <- clinical_data |>
  mutate(
    DGEA_HER2 = case_when(
      pam50 == "HER2-enriched" ~ 1, 
      TRUE ~ 0
    ), 
    DGEA_Luminal_A = case_when(
      pam50 == "Luminal A" ~ 1, 
      TRUE ~ 0
  ), 
    DGEA_Luminal_B = case_when(
      pam50 == "Luminal B" ~ 1, 
      TRUE ~ 0
      ),
    DGEA_Basal_like = case_when(
      pam50 == "Basal-like" ~ 1, 
      TRUE ~ 0)) |> 
  mutate(er_status = as.factor(er_status),  
    er_status = fct_recode(er_status, "0" = "Negative", "1" = "Positive"),  # Recode
    er_status = as.numeric(as.character(er_status))) |> 
  mutate(pr_status = as.factor(pr_status),  
    pr_status = fct_recode(pr_status, "0" = "Negative", "1" = "Positive"),  # Recode
    pr_status = as.numeric(as.character(pr_status))) |> 
  mutate(her2_final_status = as.factor(her2_final_status),  
    her2_final_status = fct_recode(her2_final_status, "0" = "Negative", "1" = "Positive"),  # Recode
    her2_final_status = as.numeric(as.character(her2_final_status)))

clinical_corr <- clinical_data |> 
  select(c("er_status","pr_status", "her2_final_status", "DGEA_HER2", "DGEA_Luminal_A", "DGEA_Luminal_B", "DGEA_Basal_like"))


corr_pearson <- clinical_corr |> 
  correlate() |> 
  focus(er_status, pr_status, her2_final_status)
  
  
corr_kendall <- clinical_corr |> 
  correlate(method = "kendall") |>
  focus(er_status, pr_status, her2_final_status)

  
```


Visualize
```{r}

#matrix <- corr_kendall |> 
#  column_to_rownames(var = "term") |> 
#  as.matrix()

#ggcorrplot(matrix, 
#           ggtheme = ggplot2::theme_light(),
#           lab = TRUE,
#           title = "Correlation matrix of tumor subclasses and her2, progesteron and estrogen status",
#           legend.title = "Correlation",
#           colors = c("blue", "white", "red"))

#print(corr_kendall)

c<- corr_kendall |> 
  dplyr::rename(subclass = term) |>  
  pivot_longer(ends_with("status"), 
               names_to = "status", 
               values_to = "corr") |> 
  mutate(
    subclass = recode(subclass, 
                      "DGEA_HER2" = "HER2", 
                      "DGEA_Luminal_A" = "Luminal A",
                      "DGEA_Luminal_B" = "Luminal B",
                      "DGEA_Basal_like" = "Basal-like"),
    status = recode(status, 
                    "er_status" = "Estrogen", 
                    "pr_status" = "Progesteron",
                    "her2_final_status" = "her2")
  ) |> 
  mutate(
    subclass = factor(subclass, levels = unique(subclass)),
    status = factor(status, levels = unique(status))
  )



  Corr_subclasses <- ggplot(c, aes(x = subclass, y = status, fill = corr)) +
  geom_tile(height=0.8, width=0.8) +
    geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  theme_minimal() +
  coord_equal() +
  labs(title = "Correlation matrix of tumor subclasses and ER, PR and her2 status", x = "Subclass", y = "Status", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 13),
        axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
        axis.text.y = element_text(size = 8, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())
  

ggsave("../plots/Corr_subclasses.png", Corr_subclasses , height = 4, width = 6)

           

```





Try corr matrix with Cramer's V

```{r}
cramers_overview <- colpair_map(clinical_corr, cramerV) |> 
  focus(er_status, pr_status, her2_final_status)

cramerscorr <- cramers_overview |> 
  dplyr::rename(subclass = term) |>  
  pivot_longer(ends_with("status"), 
               names_to = "status", 
               values_to = "corr") |> 
  mutate(
    subclass = recode(subclass, 
                      "DGEA_HER2" = "HER2", 
                      "DGEA_Luminal_A" = "Luminal A",
                      "DGEA_Luminal_B" = "Luminal B",
                      "DGEA_Basal_like" = "Basal-like"),
    status = recode(status, 
                    "er_status" = "Estrogen", 
                    "pr_status" = "Progesteron",
                    "her2_final_status" = "her2")
  ) |> 
  mutate(
    subclass = factor(subclass, levels = unique(subclass)),
    status = factor(status, levels = unique(status))
  )



  ggplot(cramerscorr, aes(x = subclass, y = status, fill = corr)) +
  geom_tile(height=0.8, width=0.8) +
    geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  theme_minimal() +
  coord_equal() +
  labs(title = "Correlation matrix of tumor subclasses and ER, PR and her2 status", x = "Subclass", y = "Status", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 12),
        axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
        axis.text.y = element_text(size = 8, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())
```




##Correlation matrix for sign. protein expression vs. HER2, PR, Er
Load differential expression data

```{r}
HER2 <- read_table("../data/diff_exp_HER2.tsv")
Luminal_A <- read_table("../data/diff_exp_Luminal_A.tsv")
Luminal_B <- read_table("../data/diff_exp_Luminal_B.tsv")
Basal_like <- read_table("../data/diff_exp_Basal_like.tsv")

```


Load and join clinical and protein data
```{r}

protein_data <- read.table("../data/count_matrix_1_77.tsv", header = T) |> 
  pivot_longer(cols = -id, names_to = 'variable', values_to = 'value') |>
  pivot_wider(names_from = id, values_from = value) |> 
  dplyr::rename(id = variable)


full_data <- clinical_data |> 
  left_join(protein_data, by = "id")
```




Extract significant proteins
```{r}
HER2_sign <- HER2 |>
  filter(adjusted_fdr < 0.05) |> 
  pull(var = 1)

Basal_like_sign <- Basal_like |>
  filter(adjusted_fdr < 0.05) |> 
  pull(var = 1)

Luminal_A_sign <- Luminal_A |>
  filter(adjusted_fdr < 0.05) |> 
  pull(var = 1)

Luminal_B_sign <- Luminal_B |>
  filter(adjusted_fdr < 0.05) |> 
  pull(var = 1)

Sign_proteins <- reduce(list(HER2_sign, Basal_like_sign, Luminal_A_sign, Luminal_B_sign), union)

length(Sign_proteins)
```



Retrieve names of sign proteins (Mapping)

```{r}
library(readr)
protein_IDs <- read_csv("/NP_to_gene_dictionary/protein_IDs.csv")
View(protein_IDs)
protein_mapping <- protein_IDs |> 
  select(RefSeq_accession_number, Protein.names, Gene.Names)


HER2_mapping <- HER2_sign |> 
  data.frame() |> 
  left_join(protein_mapping, by = c("HER2_sign" = "RefSeq_accession_number"))

Basal_like_mapping <- Basal_like_sign |> 
  data.frame() |> 
  left_join(protein_mapping, by = c("Basal_like_sign" = "RefSeq_accession_number"))

Luminal_A_mapping <- Luminal_A_sign |> 
  data.frame() |> 
  left_join(protein_mapping, by = c("Luminal_A_sign" = "RefSeq_accession_number"))

Luminal_B_mapping <- Luminal_B_sign |> 
  data.frame() |> 
  left_join(protein_mapping, by = c("Luminal_B_sign" = "RefSeq_accession_number"))
```



```{r}
full_data <- full_data |>
  select(c(id, er_status, pr_status, her2_final_status, DGEA_HER2, DGEA_Luminal_A, DGEA_Luminal_B, DGEA_Basal_like, all_of(Sign_proteins))) |>
  drop_na()


Sign_protein_corr <- correlate(full_data, method = "kendall") |>
  focus(er_status, pr_status, her2_final_status, DGEA_HER2, DGEA_Luminal_A, DGEA_Luminal_B, DGEA_Basal_like)


```





Visualize
```{r}

c2 <- Sign_protein_corr |> 
  dplyr::rename(Protein = term) |>   
  pivot_longer(cols = c(ends_with("status"),starts_with("DGEA")),
               names_to = "category", 
               values_to = "corr") |> 
  mutate(
    category = recode(category, 
                      "DGEA_HER2" = "HER2", 
                      "DGEA_Luminal_A" = "Luminal A",
                      "DGEA_Luminal_B" = "Luminal B",
                      "DGEA_Basal_like" = "Basal-like",
                      "er_status" = "Estrogen", 
                      "pr_status" = "Progesteron",
                      "her2_final_status" = "her2")
  ) |> 
  mutate(
    category = factor(category, levels = unique(category)))






```


Plot corr matrix with proteins
```{r}
  c2plot <- ggplot(c2, aes(x = category, y = Protein, fill = corr)) +
  geom_tile() +
    #geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  theme_minimal() +
  labs(x = "category", y = "Protein", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 5),
        axis.text.x = element_text(size = 3, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
       axis.text.y = element_text(size = 3, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())

ggsave("../plots/corr_mat_ful.png", c2plot, height = 20, width = 4)
```






Create long protein table with ids and expressions
```{r}
protein_long <- read.table("../data/count_matrix_1_77.tsv", header = T) |> 
pivot_longer(cols = -id, names_to = 'Proteins', values_to = 'Expression') 
```





##Create separate corr matrices

Her2 class
```{r}
Her2_prot <- full_data |>
  select(c(id, er_status, pr_status, her2_final_status, HER2_sign)) |> 
  drop_na()


Her2_corr <- correlate(Her2_prot, method = "kendall") |>
  focus(er_status, pr_status, her2_final_status)

Her2_corr <- Her2_corr |> 
  dplyr::rename(Protein = term) |>   
  pivot_longer(cols = ends_with("status"),
               names_to = "status", 
               values_to = "corr") |> 
  mutate(
    status = recode(status, 
                      "er_status" = "Estrogen", 
                      "pr_status" = "Progesteron",
                      "her2_final_status" = "her2")
  ) |> 
  mutate(
    status = factor(status, levels = unique(status)))

Her2_plot <- ggplot(Her2_corr, aes(x = status, y = Protein, fill = corr)) +
  geom_tile(height=0.8, width=0.8) +
    geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  scale_fill_gradient2(low="blue", mid="white", high="red", limits = c(-1,1)) +
  theme_minimal() +
  coord_equal() +
  labs(title = "HER2-positive", x = "Status", y = "Protein", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 12),
        axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
        axis.text.y = element_text(size = 8, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())


Her2_plot


ggsave("../plots/corr_Her2.png", Her2_plot, height = 6, width = 5)

```




Luminal A
```{r}
length(Luminal_A_sign)
Luminal_A_sign[c(1:5)]

Luminal_A_prot <- full_data |>
  select(c(id, er_status, pr_status, her2_final_status, Luminal_A_sign[c(1:5)])) |>   drop_na()


Luminal_A_corr <- correlate(Luminal_A_prot, method = "kendall") |>
  focus(er_status, pr_status, her2_final_status)

Luminal_A_corr <- Luminal_A_corr |> 
  dplyr::rename(Protein = term) |>   
  pivot_longer(cols = ends_with("status"),
               names_to = "status", 
               values_to = "corr") |> 
  mutate(
    status = recode(status, 
                      "er_status" = "Estrogen", 
                      "pr_status" = "Progesteron",
                      "her2_final_status" = "her2")
  ) |> 
  mutate(
    status = factor(status, levels = unique(status)))

Luminal_A_plot <- ggplot(Luminal_A_corr, aes(x = status, y = Protein, fill = corr)) +
  geom_tile(height = 0.8, width = 0.8) +
    geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  scale_fill_gradient2(low="blue", mid="white", high="red", limits = c(-1,1)) +
  theme_minimal() +
  coord_equal() +
  labs(title = "Luminal A", x = "Status", y = "Protein", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 12),
        plot.subtitle = element_text(hjust = 0.4),
        axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
        axis.text.y = element_text(size = 8, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())


Luminal_A_plot


ggsave("../plots/corr_LumA.png", Luminal_A_plot, height = 5, width = 8)

```



Luminal B
```{r}
length(Luminal_B_sign)

Luminal_B_prot <- full_data |>
  select(c(id, er_status, pr_status, her2_final_status, Luminal_B_sign)) |>   drop_na()


Luminal_B_corr <- correlate(Luminal_B_prot, method = "kendall") |>
  focus(er_status, pr_status, her2_final_status)

Luminal_B_corr <- Luminal_B_corr |> 
  dplyr::rename(Protein = term) |>   
  pivot_longer(cols = ends_with("status"),
               names_to = "status", 
               values_to = "corr") |> 
  mutate(
    status = recode(status, 
                      "er_status" = "Estrogen", 
                      "pr_status" = "Progesteron",
                      "her2_final_status" = "her2")
  ) |> 
  mutate(
    status = factor(status, levels = unique(status)))

Luminal_B_plot <- ggplot(Luminal_B_corr, aes(x = status, y = Protein, fill = corr)) +
  geom_tile(height = 0.8, width = 0.8) +
    geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  scale_fill_gradient2(low="blue", mid="white", high="red", limits = c(-1,1)) +
  theme_minimal() +
  coord_equal()+
  labs(title = "Luminal B", x = "Status", y = "Protein", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 12),
        plot.subtitle = element_text(hjust = 0.4),
        axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
        axis.text.y = element_text(size = 8, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())


Luminal_B_plot


ggsave("../plots/corr_LumB.png", Luminal_B_plot, height = 5, width = 8)

```




Basal-like
```{r}
length(Basal_like_sign)
Basal_like_sign[c(1:5)]

Basal_like_prot <- full_data |>
  select(c(id, er_status, pr_status, her2_final_status, Basal_like_sign[c(1:5)])) |>   drop_na()


Basal_like_corr <- correlate(Basal_like_prot, method = "kendall") |>
  focus(er_status, pr_status, her2_final_status)

Basal_like_corr <- Basal_like_corr |> 
  dplyr::rename(Protein = term) |>   
  pivot_longer(cols = ends_with("status"),
               names_to = "status", 
               values_to = "corr") |> 
  mutate(
    status = recode(status, 
                      "er_status" = "Estrogen", 
                      "pr_status" = "Progesteron",
                      "her2_final_status" = "her2")
  ) |> 
  mutate(
    status = factor(status, levels = unique(status)))

Basal_like_plot <- ggplot(Basal_like_corr, aes(x = status, y = Protein, fill = corr)) +
  geom_tile(height = 0.8, width = 0.8) +
    geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  #scale_fill_gradientn(colours=c( "blue",  
  #                               "white",
   #                              "red"),
    #                   values=(c( -1,
     ##                                  1)),
       #                guide="colorbar")+
  scale_fill_gradient2(midpoint = 0, low="blue", mid="white", high="red", limits = c(-1,1)) +
  theme_minimal() +
  coord_equal()+
  labs(title = "Basal-like", x = "Status", y = "Protein", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 12),
        plot.subtitle = element_text(hjust = 0.4),
        axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
        axis.text.y = element_text(size = 8, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())


Basal_like_plot


ggsave("../plots/corr_Basal_like.png", Basal_like_plot, height = 8, width = 6)

```



```{r}
p4 <-(Basal_like_plot + Her2_plot) / (Luminal_A_plot + Luminal_B_plot)+
  plot_layout(guides = 'collect') +
  plot_annotation(title = "Correlation of PR, ER and her2 status with significant proteins", 
        subtitle = "Using top 5 most significant proteins for each PAM50 classification")&
  theme(
    plot.title = element_text(size = 16, hjust = 0.5), # Adjust title size and horizontal position
    plot.subtitle = element_text(size = 14, hjust = 0.5) # Adjust subtitle size and horizontal position
  )

ggsave("../plots/Protein_corr_4p.png", p4 , height = 8, width = 10)
```










Do correlation with Cramers V
```{r}
cramers_protein_corr <- colpair_map(full_data,cramerV) |> 
  focus(er_status, pr_status, her2_final_status, DGEA_HER2, DGEA_Luminal_A, DGEA_Luminal_B, DGEA_Basal_like) |> 
  slice(-c(1))



cramers_protein_corr <- cramers_protein_corr |> 
  dplyr::rename(Protein = term) |>   
  pivot_longer(cols = c(ends_with("status"),starts_with("DGEA")),
               names_to = "category", 
               values_to = "corr") |> 
  mutate(
    category = recode(category, 
                      "DGEA_HER2" = "HER2", 
                      "DGEA_Luminal_A" = "Luminal A",
                      "DGEA_Luminal_B" = "Luminal B",
                      "DGEA_Basal_like" = "Basal-like",
                      "er_status" = "Estrogen", 
                      "pr_status" = "Progesteron",
                      "her2_final_status" = "her2")
  ) |> 
  mutate(
    category = factor(category, levels = unique(category)))

  
```



##Create separate corr matrices with Cramers V

Her2 class
```{r}



Her2_cramer <- colpair_map(Her2_prot,cramerV) |> 
  focus(er_status, pr_status, her2_final_status) |> 
  slice(-c(1))


Her2_cramer <- Her2_cramer |> 
  dplyr::rename(Protein = term) |>   
  pivot_longer(cols = ends_with("status"),
               names_to = "status", 
               values_to = "corr") |> 
  mutate(
    status = recode(status, 
                      "er_status" = "Estrogen", 
                      "pr_status" = "Progesteron",
                      "her2_final_status" = "her2")
  ) |> 
  mutate(
    status = factor(status, levels = unique(status)))

Her2_cramerplot <- ggplot(Her2_cramer, aes(x = status, y = Protein, fill = corr)) +
  geom_tile(height=0.8, width=0.8) +
    geom_text(aes(label = sprintf("%.2f", corr)), color = "black", size = 3)+
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  theme_minimal() +
  coord_equal() +
  labs(title = "HER2-positive", subtitle = "Correlation of status and protein expression" , x = "Status", y = "Protein", fill = "Correlation") +
  theme(plot.title = element_text(hjust = 0.4, size = 12),
        axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1, margin = margin(-3, 0, 0, 0)),
        axis.text.y = element_text(size = 8, margin = margin(0, -3, 0, 0)),
        panel.grid.major = element_blank())


Her2_cramerplot


ggsave("../plots/corr_Her2.png", Her2_plot, height = 6, width = 5)

```
